---
title: "R Notebook"
author: Amol
date: 4/5/2020
output: html_notebook
---

The [R plugin](https://www.jetbrains.com/help/pycharm/r-plugin-support.html) for IntelliJ-based IDEs provides
handy capabilities to work with the [R Markdown](https://www.jetbrains.com/help/pycharm/r-markdown.html) files.
To [add](https://www.jetbrains.com/help/pycharm/r-markdown.html#add-code-chunk) a new R chunk,
position the caret at any line or the code chunk, then click "+".

The code chunk appears:
```{r}
```

Type any R code in the chunk, for example:
```{r, setup}
set.seed(42)

load("churn.Rdata")
install.packages("caret")
library("caret")
install.packages("C50")
library("C50")
install.packages("glmnet")
library("glmnet")
install.packages("ranger")
library("ranger")
install.packages("Metrics")
library("Metrics")

```

```{r, Importing files}

grades<- read.csv("grade.csv")

head(grades)
```
```{r, cretion of traning validation and test datasets}
# Look at the data
str(grades)

# Set seed and create assignment
#creation of dataset to create regression tree.
set.seed(1)
assignment <- sample(1:3, size = nrow(grades), prob = c(0.7,0.15,0.15), replace = TRUE)

# Create a train, validation and tests from the original data frame
grade_train <- grades[assignment == 1, ]    # subset grade to training indices only
grade_valid <- grades[assignment == 2, ]  # subset grade to validation indices only
grade_test <- grades[assignment == 3, ]   # subset grade to test indices only
```
```{r, train model visulizing model  }
# Train the model
grade_model <- rpart(formula = final_grade ~ .,
                     data = grade_train,
                     method = "anova")

# Look at the model output
print(grade_model)

# Plot the tree model
rpart.plot(x = grade_model, yesno = 2, type = 0, extra = 0)


```
```{r, calculation of RMSE}
# Generate predictions on a test set
pred <- predict(object = grade_model,   # model object
                newdata = grade_test)  # test dataset

# Compute the RMSE
Metrics::rmse(grade_test$final_grade, pred)

```


```{r, plotting cp}
#plot cp
plotcp(grade_model)

# Print the "CP Table"
print(grade_model$cptable)

# Retrieve optimal cp value based on cross-validated error
opt_index <- which.min(grade_model$cptable[, "xerror"])
cp_opt <- grade_model$cptable[opt_index, "CP"]

# Prune the model (to optimized cp value)
grade_model_opt <- prune(tree = grade_model,
                         cp = cp_opt)

# Plot the optimized model
rpart.plot(x = grade_model_opt, yesno = 2, type = 0, extra = 0)
rpart.plot(x = grade_model, yesno = 2, type = 0, extra = 0)

```
```{r, grid search manual}
# Establish a list of possible values for minsplit and maxdepth
#manually specifying gridsearch.
minsplit <- seq(1, 4, 1)
maxdepth <- seq(1, 6, 1)

# Create a data frame containing all combinations
hyper_grid <- expand.grid(minsplit = minsplit, maxdepth = maxdepth)

# Check out the grid
head(hyper_grid)

# Print the number of grid combinations
nrow(hyper_grid)


# Number of potential models in the grid
num_model <- nrow(hyper_grid)

# Create an empty list to store models
grade_models <- list()

# Write a loop over the rows of hyper_grid to train the grid of models
for (i in 1:num_models) {

  # Get minsplit, maxdepth values at row i
  minsplit <- hyper_grid$minsplit[i]
  maxdepth <- hyper_grid$maxdepth[i]

  # Train a model and store in the list
  grade_models[[i]] <- rpart(formula = final_grade ~ .,
                             data = grade_train,
                             method = "anova",
                             minsplit = minsplit,
                             maxdepth = maxdepth)
}


```

```{r}
# Number of potential models in the grid
num_models <- length(grade_models)

# Create an empty vector to store RMSE values
rmse_values <- c()

# Write a loop over the models to compute validation RMSE
for (i in 1:num_models) {

  # Retrieve the i^th model from the list
  model <- grade_models[[i]]

  # Generate predictions on grade_valid
  pred <- predict(object = model,
                  newdata = grade_valid)

  # Compute validation RMSE and add to the
  rmse_values[i] <- rmse(actual = grade_valid$final_grade,
                         predicted = pred)
}

# Identify the model with smallest validation set RMSE
best_model <- grade_models[[which.min(rmse_values)]]

# Print the model paramters of the best model
best_model$control

# Compute test set RMSE on best_model
pred <- predict(object = best_model,
                newdata = grade_test)
rmse(actual = grade_test$final_grade,
     predicted = pred)
```

Now, click the **Run** button on the chunk toolbar to [execute](https://www.jetbrains.com/help/pycharm/r-markdown.html#run-r-code) the chunk code. The result should be placed under the chunk.
Click the **Knit and Open Document** to built and preview an output.